---
title: "**R** HIGHS Interface"
author: "Florian Schwendinger</br>"
date: "Updated: `r Sys.Date()`"
output:
  github_document:
    number_sections: true
---

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/highs)](https://CRAN.R-project.org/package=highs)
[![Licence](https://img.shields.io/cran/l/highs)](https://www.gnu.org/licenses/gpl-2.0.en.html)
<!-- badges: end -->

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This repository contains an **R** interface to the [**HiGHS**](https://github.com/ERGO-Code/HiGHS) solver.
The [**HiGHS**](https://github.com/ERGO-Code/HiGHS) solver,
is a **high**-performance open-source **solver** for
solving linear programming (LP), mixed-integer programming
(MIP) and quadratic programming (QP) optimization problems.


# Installation
The package can be installed from [**CRAN**](https://CRAN.R-project.org/package=highs)
```{r, eval = FALSE}
install.packages("highs")
```
or [**GitLab**](https://gitlab.com/roigrp/solver/highs).
```{r, eval = FALSE}
remotes::install_gitlab("roigrp/solver/highs")
```

### Using a preinstalled HiGHS library
It is possible to use a precompile HiGHS library by providing
the system variable `R_HIGHS_LIB_DIR`.
For example I used 
```sh
mkdir build
cd build
cmake .. -DCMAKE_INSTALL_PREFIX=/Z/bin/highslib -DCMAKE_POSITION_INDEPENDENT_CODE:bool=ON -DSHARED:bool=OFF -DBUILD_TESTING:bool=OFF
make install
```
to install the **HiGHS** library to `/Z/bin/highslib`

```{r, eval = FALSE}
Sys.setenv(R_HIGHS_LIB_DIR = "/Z/bin/highslib")
install.packages("highs")
# or 
# remotes::install_gitlab("roigrp/solver/highs")
```


# Basic usage
```{r}
library("highs")

args(highs_solve)
```


## LP
```{r}
# Minimize
#  x_0 +  x_1 + 3
# Subject to
#                 x_1 <= 7
#  5 <=   x_0 + 2 x_1 <= 15
#  6 <= 3 x_0 + 2 x_1
#  0 <=   x_0         <= 4
#  1 <=           x_1
A <- rbind(c(0, 1), c(1, 2), c(3, 2))
s <- highs_solve(L = c(1.0, 1), lower = c(0, 1), upper = c(4, Inf),
                 A = A, lhs = c(-Inf, 5, 6), rhs = c(7, 15, Inf),
                 offset = 3)
str(s)
```


## QP
```{r}
# Minimize
#  0.5 x^2 - 2 x + y
# Subject to
#  x <= 3
zero <- .Machine$double.eps * 100
Q <- rbind(c(1, 0), c(0, zero))
L <- c(-2, 1)
A <- t(c(1, 0))

cntrl <- list(log_dev_level = 0L)
s <- highs_solve(Q = Q, L = L, A = A, lhs = 0, rhs = 3, control = cntrl)
str(s)
```


# Options
The function `highs_available_solver_options` lists the available solver options
```{r}
d <- highs_available_solver_options()
d[["option"]] <- sprintf("`%s`", d[["option"]])
knitr::kable(d, row.names = FALSE)
```

for additional information see the [HiGHS homepage](https://highs.dev/).	


# Status codes
HiGHS currently has the following status codes defined in `HConst.h"`.

| enumerator               | status | message                            |
| -------------------------|-------:|------------------------------------|
| `kNotset`                |      0 | `"Not Set"`                        |
| `kLoadError`             |      1 | `"Load error"`                     |
| `kModelError`            |      2 | `"Model error"`                    |
| `kPresolveError`         |      3 | `"Presolve error"`                 |
| `kSolveError`            |      4 | `"Solve error"`                    |
| `kPostsolveError`        |      5 | `"Postsolve error"`                |
| `kModelEmpty`            |      6 | `"Empty"`                          |
| `kOptimal`               |      7 | `"Optimal"`                        |
| `kInfeasible`            |      8 | `"Infeasible"`                     |
| `kUnboundedOrInfeasible` |      9 | `"Primal infeasible or unbounded"` |
| `kUnbounded`             |     10 | `"Unbounded"`                      |
| `kObjectiveBound`        |     11 | `"Bound on objective reached"`     |
| `kObjectiveTarget`       |     12 | `"Target for objective reached"`   |
| `kTimeLimit`             |     13 | `"Time limit reached"`             |
| `kIterationLimit`        |     14 | `"Iteration limit reached"`        |
| `kUnknown`               |     15 | `"Unknown"`                        |
| `kMin`                   |      0 | `"Not Set"`                        |
| `kMax`                   |     15 | `"Unknown"`                        |


