---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# **R** HIGHS Interface

This repository contains an **R** interface to the **HiGHS** solver.
The [HiGHS](https://github.com/ERGO-Code/HiGHS) solver,
is a **high**-performance open-source **solver** for
solving linear programming (LP), mixed-integer programming
(MIP) and quadratic programming (QP) optimization problems.


## Installation
```{r, eval = FALSE}
remotes::install_gitlab("roigrp/solver/highs")
```


## Basic usage
```{r}
library("highs")

args(highs_solve)
```


### LP
```{r}
# Minimize
#  x_0 +  x_1 + 3
# Subject to
#                 x_1 <= 7
#  5 <=   x_0 + 2 x_1 <= 15
#  6 <= 3 x_0 + 2 x_1
#  0 <=   x_0         <= 4
#  1 <=           x_1
A <- rbind(c(0, 1), c(1, 2), c(3, 2))
s <- highs_solve(L = c(1.0, 1), lower = c(0, 1), upper = c(4, Inf),
                 A = A, lhs = c(-Inf, 5, 6), rhs = c(7, 15, Inf),
                 offset = 3)
str(s)
```


## QP
```{r}
# Minimize
#  0.5 x^2 - 2 x + y
# Subject to
#  x <= 3
zero <- .Machine$double.eps * 100
Q <- rbind(c(1, 0), c(0, zero))
L <- c(-2, 1)
A <- t(c(1, 0))

cntrl <- list(log_dev_level = 0L)
s <- highs_solve(Q = Q, L = L, A = A, lhs = 0, rhs = 3, control = cntrl)
str(s)
```


## Options
The function `highs_available_solver_options` lists the available solver options
```{r}
highs_available_solver_options()
```

for additional information see the [HiGHS homepage](https://highs.dev/).	

