---
title: "**R** HIGHS Interface"
author: "Florian Schwendinger</br>"
date: "Updated: `r Sys.Date()`"
output:
  github_document:
    number_sections: true
---

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/highs)](https://CRAN.R-project.org/package=highs)
[![Licence](https://img.shields.io/cran/l/highs)](https://www.gnu.org/licenses/gpl-2.0.en.html)
<!-- badges: end -->

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This repository contains an **R** interface to the [**HiGHS**](https://github.com/ERGO-Code/HiGHS) solver.
The [**HiGHS**](https://github.com/ERGO-Code/HiGHS) solver,
is a **high**-performance open-source **solver** for
solving linear programming (LP), mixed-integer programming
(MIP) and quadratic programming (QP) optimization problems.


# Installation

## Install system dependencies

### Debian/Ubuntu
```sh
sudo apt-get install libhighs-dev
```

### Fedora
```sh
sudo dnf install coin-or-HiGHS-devel
```

### Arch Linux
```sh
sudo pacman -S highs
```

### Windows    
- Install Rtools
- Install highs via MSYS2 
There is there is more than one version of highs available in MSYS2,
you most likely need the following
```sh
pacman -S mingw-w64-x86_64-highs
```
a full list can be found at `https://packages.msys2.org/base/mingw-w64-highs`.

## Install R package

The package can be installed from [**CRAN**](https://CRAN.R-project.org/package=highs)
```{r, eval = FALSE}
install.packages("highs")
```
or [**GitLab**](https://gitlab.com/roigrp/solver/highs).
```{r, eval = FALSE}
remotes::install_gitlab("roigrp/solver/highs")
```


# Package
The **highs** package provides an API similar to **Rglpk** and a low level API to the **HiGHS** solver.
For most users using `highs_solve` as shown below should be the best choice.

The package functions can be grouped into the following categories:

1. The main function **`highs_solve`**.
2. Object style wrappers for the model and the solver **`highs_model`** and **`highs_solver`**.
3. Function **`highs_control`** to construct the control object for **`highs_solve`**and **`highs_solver`**.
4. Low-level API wrapper functions to create and modify models **`hi_new_model`**
   and other functions starting with **`hi_model_`**.
5. Low-level API wrapper functions to create and modify solvers **`hi_new_solver`**
   and other functions starting with **`hi_solver_`**.
6. Functions to create example models **`example_model`** and solvers **`example_solver`** for the documentation examples.
7. Function **`highs_available_solver_options`** to get the available solver options.
8. Function **`highs_write_model`** to write the model to a file.

```{r}
library("highs")
```

## Examples
The the example models and solvers are included to have small examples available for the manual.

```{r}
writeLines(ls("package:highs", pattern = "^example"))
```

## Low level model functions
The low-level model functions allow to create and modify models.
More details and examples can be found in the manual.

```{r}
writeLines(ls("package:highs", pattern = "^hi(|_new)_model"))
```

## Low level solver functions
The low-level solver functions allow to create and modify solvers.
More details and examples can be found in the manual.

```{r}
writeLines(ls("package:highs", pattern = "^hi(|_new)_solver"))
```

## High level functions
The high level functions allow to work with models and solvers.
More details and examples can be found in the manual.

```{r}
args(highs_model)
args(highs_solver)
args(highs_control)
args(highs_write_model)
```

## Main function
The main function `highs_solve`.

```{r}
library("highs")

args(highs_solve)
```


## LP
```{r}
# Minimize
#  x_0 +  x_1 + 3
# Subject to
#                 x_1 <= 7
#  5 <=   x_0 + 2 x_1 <= 15
#  6 <= 3 x_0 + 2 x_1
#  0 <=   x_0         <= 4
#  1 <=           x_1
A <- rbind(c(0, 1), c(1, 2), c(3, 2))
s <- highs_solve(L = c(1.0, 1), lower = c(0, 1), upper = c(4, Inf),
                 A = A, lhs = c(-Inf, 5, 6), rhs = c(7, 15, Inf),
                 offset = 3)
str(s)
```


## QP
```{r}
# Minimize
#  0.5 x^2 - 2 x + y
# Subject to
#  x <= 3
zero <- .Machine$double.eps * 100
Q <- rbind(c(1, 0), c(0, zero))
L <- c(-2, 1)
A <- t(c(1, 0))

cntrl <- list(log_dev_level = 0L)
s <- highs_solve(Q = Q, L = L, A = A, lhs = 0, rhs = 3, control = cntrl)
str(s)
```

# Additional information
## Sparse matrices
The **HiGHs** **C++** library internally supports the matrix formats 
csc (compressed sparse column matrix) and csr (compressed Sparse Row array).
The **highs** package currently supports the following matrix classes:

1. `"matrix"` dense matrices,   
2. `"dgCMatrix"` compressed sparse column matrix from the **Matrix** package,   
3. `"dgRMatrix"` compressed sparse row matrix from the **Matrix** package,   
4. `"matrix.csc"` compressed sparse column matrix from the **SparseM** package,   
5. `"matrix.csr"` compressed sparse row matrix from the **SparseM** package,   
6. `"simple_triplet_matrix"` coordinate format from the **slam** package.    

If the constraint matrix `A` is provided as `dgCMatrix`, `dgRMatrix`, `matrix.csc`
or `matrix.csr` the underlying data is directly passed to **HiGHs** otherwise
it is first transformed into the csc format an afterwards passed to **HiGHs** 

```{r, eval = FALSE}
library("Matrix")

A <- rbind(c(0, 1), c(1, 2), c(3, 2))
csc <- as(A, "CsparseMatrix")  # dgCMatrix
s0 <- highs_solve(L = c(1.0, 1), lower = c(0, 1), upper = c(4, Inf),
                  A = csc, lhs = c(-Inf, 5, 6), rhs = c(7, 15, Inf),
                  offset = 3)

csr <- as(A, "RsparseMatrix")  # dgRMatrix
s1 <- highs_solve(L = c(1.0, 1), lower = c(0, 1), upper = c(4, Inf),
                  A = csr, lhs = c(-Inf, 5, 6), rhs = c(7, 15, Inf),
                  offset = 3)

library("SparseM")

csc <- as.matrix.csc(A)
s2 <- highs_solve(L = c(1.0, 1), lower = c(0, 1), upper = c(4, Inf),
                  A = csc, lhs = c(-Inf, 5, 6), rhs = c(7, 15, Inf),
                  offset = 3)

csr <- as.matrix.csr(A)
s3 <- highs_solve(L = c(1.0, 1), lower = c(0, 1), upper = c(4, Inf),
                  A = csr, lhs = c(-Inf, 5, 6), rhs = c(7, 15, Inf),
                  offset = 3)

library("slam")
stm <- as.simple_triplet_matrix(A)
s4 <- highs_solve(L = c(1.0, 1), lower = c(0, 1), upper = c(4, Inf),
                  A = stm, lhs = c(-Inf, 5, 6), rhs = c(7, 15, Inf),
                  offset = 3)
```


# Options
The function `highs_available_solver_options` lists the available solver options
```{r}
d <- highs_available_solver_options()
d[["option"]] <- sprintf("`%s`", d[["option"]])
knitr::kable(d, row.names = FALSE)
```

for additional information see the [HiGHS homepage](https://highs.dev/).	


# Status codes
HiGHS currently has the following status codes defined in `HConst.h"`.

| enumerator               | status | message                            |
| -------------------------|-------:|------------------------------------|
| `kNotset`                |      0 | `"Not Set"`                        |
| `kLoadError`             |      1 | `"Load error"`                     |
| `kModelError`            |      2 | `"Model error"`                    |
| `kPresolveError`         |      3 | `"Presolve error"`                 |
| `kSolveError`            |      4 | `"Solve error"`                    |
| `kPostsolveError`        |      5 | `"Postsolve error"`                |
| `kModelEmpty`            |      6 | `"Empty"`                          |
| `kOptimal`               |      7 | `"Optimal"`                        |
| `kInfeasible`            |      8 | `"Infeasible"`                     |
| `kUnboundedOrInfeasible` |      9 | `"Primal infeasible or unbounded"` |
| `kUnbounded`             |     10 | `"Unbounded"`                      |
| `kObjectiveBound`        |     11 | `"Bound on objective reached"`     |
| `kObjectiveTarget`       |     12 | `"Target for objective reached"`   |
| `kTimeLimit`             |     13 | `"Time limit reached"`             |
| `kIterationLimit`        |     14 | `"Iteration limit reached"`        |
| `kUnknown`               |     15 | `"Unknown"`                        |
| `kMin`                   |      0 | `"Not Set"`                        |
| `kMax`                   |     15 | `"Unknown"`                        |


